---
output:
  word_document: default
  html_document: default
---
# Module 3 - Assignment 3

## Simis, Kathleen

### More Practice with Plots

```{r}
library(tidyverse)
GOOG <- read_csv("GOOG.csv")
```

#### Google Stock Price Plots

The following is an analysis of Google’s stock price from October of 2019 to October of 2020. This will include a scatter, line, bar, histogram and boxplot. All the plots represent the closing price on the dates listed on the x-axis.

```{r Basic Plots}
ggplot(data=GOOG , aes(x=Date, y=Close)) +
  geom_point()
ggplot(data=GOOG, aes(x=Date, y=Close)) +
  geom_line()
ggplot(data=GOOG, aes(x=Date, y=Close)) +
  geom_col()
ggplot(data=GOOG, aes(Close)) +
  geom_histogram()
ggplot(data=GOOG, aes(Close)) +
  geom_boxplot()
ggplot(data=GOOG, aes(x = Date, y = Close)) +
  geom_line() +
  geom_point()
```

Of all of the above plots, I feel that the line graph is the easiest to read. While similar information can be seen in the scatterplot and in the line graph with dots, I don't feel like the dots are neccessary, and the simplicity of the line is better able to capture the general change over time. 

#### Google Stock Daily Price Comparisons

```{r Advanced Plots}
ggplot(GOOG) +
  geom_line(aes(x=Date, y=High, color="High")) +
  geom_line(aes(x=Date, y=Low, color="Low"))
ggplot(GOOG) +
  geom_line(aes(x=Date, y=Open, color="Open")) +
  geom_line(aes(x=Date, y=Close, color="Close"))
ggplot(GOOG, aes(x=Date, y=Close)) +
  geom_area(color="black",fill="blue",alpha=.2)
ggplot(GOOG, aes(x=Date, y=Open)) +
  geom_area(color="blue",fill="green",alpha=.1)
MayPrice <- subset(GOOG, Date > "2020-04-30" & Date < "2020-06-01")
ggplot(MayPrice, aes(x=Date, y=Close)) +
  geom_point() +
  geom_text(aes(label=round(Close)), size=3, vjust=-1)
AprilPrice <- subset(GOOG, Date > "2020-03-31" & Date < "2020-05-01")
ggplot(AprilPrice, aes(x=Date, y=Open)) +
  geom_point() +
  geom_text(aes(label=round(Open)), size=2, vjust=-1.5)
```



