---
output:
  word_document: default
  html_document: default
---
# Module 4 - Assignment 1

## Simis, Kathleen

### Data Transformation

```{r}
library(tidyverse)
state_income <- read_csv("state_income.csv")
```

#### State Incomes

The first thing I will be doing is creating a subset of data from the larger state_income.csv file. The subset will focus on the following variables: State_Name, State_ab (the abbreviated state name reported by the U.S. Census Bureau for the specified geographic location), County, City, Type (the place Type reported by the U.S. Census Bureau for the specified geographic location), ALand (the Square area of land at the geographic or track location), Mean (mean household income of the specified geographic location), Median (the median household income of the specified geographic location), and Stdev (the standard deviation of the household income for the specified geographic
location).

```{r}
state_income2 <- select(state_income, State_Name, State_ab, County, City, Type, ALand, Mean, Median, Stdev)
state_income2 <- select(state_income2, State_ab, everything())
head(state_income2)
```

```{r}
state_income2 <- rename(state_income2, SquareArea = ALand)
state_income2 <- rename(state_income2, IncomeMean = Mean)
state_income2 <- rename(state_income2, IncomeMedian = Median)
state_income2 <- rename(state_income2, IncomeStDev = Stdev)
head(state_income2)
```
```{r}
NC_income <- filter (state_income2, State_Name == "North Carolina")
head(NC_income)
```

#### NC Incomes

I will be using the NC_income dataset to create summaries of the incomes within North Carolina. This will include summaries by county, city and type.

```{r}
NC_income <- arrange(NC_income, County)
head(NC_income)
```

```{r}
summary1 <- group_by(NC_income,County)
summary1 <- summarise (summary1, mean= mean(IncomeMean))
summary2 <- NC_income %>%
  group_by(City) %>%
  summarise (mean = mean(IncomeMean))
```

The above code creates two summary sets, one displaying summary statistics related to the mean income by county, and one displaying summary statistics related to the mean income by city. 

```{r}
summary3 <- NC_income %>%
  group_by(Type) %>%
  summarise (mean = mean(IncomeMean))
```

#### Income Visualization

```{r}
ggplot(data=summary1, aes(x=County,y=mean)) +
  geom_point() +
  labs(y='Income') +
  theme(axis.text.x = element_text(angle = 90,vjust =0.5, hjust=1))
```

Per the scatterplot above, we can see that Wake County has the highest average income, while Stokes county has the lowest average income, followed closely by Surry County. 

```{r}
AvgStateIncome <- state_income2 %>%
  group_by(State_ab) %>%
  summarise (mean = mean(IncomeMean))
AvgStateIncome <- filter(AvgStateIncome, grepl("N",State_ab))
ggplot(data=AvgStateIncome, aes(x=State_ab, y=mean)) +
  geom_point() +
  labs(x= 'State Abbreviation', y='Average Income')
```

Based on the above scatterplot, we can see that New Jersey has the highest average income, while Tennessee has the lowest. 